---
title: "team selection analysis"
output: html_document
---

```{r library}
library(readxl)
library(dplyr)
```

Load data
```{r}
file_path <- "/Users/liushuhao/Desktop/SOA Case Competitoin/data/2022-student-research-case-study-player-data_modified.xlsx"

#read in tables
LeagueShoot <- read_excel(file_path, sheet = "League Shooting", skip = 11,col_names =T, na = "N/A")
LeaguePass <- read_excel(file_path, sheet = "League Passing", skip = 11,col_names =T)
LeagueDefense <- read_excel(file_path, sheet = "League Defense", skip = 11,col_names =T, na = "N/A")
TournShoot <- read_excel(file_path, sheet = "Tournament Shooting", skip = 11,col_names =T, na = "N/A")
TournPass <- read_excel(file_path, sheet = "Tournament Passing", skip = 11,col_names =T)
TournDefense <- read_excel(file_path, sheet = "Tournament Defense", skip = 11,col_names =T, na = "N/A")
tables <- list("LeagueShoot" = LeagueShoot, "LeaguePass" = LeaguePass, "LeagueDefense" = LeagueDefense, "TournShoot" = TournShoot, "TournPass" = TournPass, "TournDefense" = TournDefense)
```


calculate %s
```{r}
LeagueShoot$'Gper90' <- LeagueShoot$Gls/LeagueShoot$'90s'
LeagueShoot$'Performance%' <- LeagueShoot$'Performance PKatt'/LeagueShoot$'Performance PK'
TournShoot$'Gper90' <- TournShoot$Gls/TournShoot$'90s'
TournShoot$'Performance%' <- TournShoot$'Performance PKatt'/TournShoot$'Performance PK'

LeagueDefense$'TklW%' <- LeagueDefense$'Tackles TklW'/LeagueDefense$'Tackles Tkl'
LeagueDefense$'Tackles Tkl/min' <- LeagueDefense$'Tackles Tkl'/LeagueDefense$'90s'
LeagueDefense$'Vs Dribbles Past/min' <- LeagueDefense$'Vs Dribbles Past'/LeagueDefense$'90s'
LeagueDefense$'Pressures Succ %' <- LeagueDefense$'Pressures Succ'/LeagueDefense$'Pressures Press'
LeagueDefense$'Blocks Blocks/min' <- LeagueDefense$'Blocks Blocks'/LeagueDefense$'90s'
LeagueDefense$'Int/min' <- LeagueDefense$'Int'/LeagueDefense$'90s'
LeagueDefense$'Clr/min' <- LeagueDefense$'Clr'/LeagueDefense$'90s'
LeagueDefense$'Err/min' <- LeagueDefense$'Err'/LeagueDefense$'90s'
TournDefense$'TklW%' <- TournDefense$'Tackles TklW'/TournDefense$'Tackles Tkl'
TournDefense$'Tackles Tkl/min' <- TournDefense$'Tackles Tkl'/TournDefense$'90s'
TournDefense$'Vs Dribbles Past/min' <- TournDefense$'Vs Dribbles Past'/TournDefense$'90s'
TournDefense$'Pressures Succ %' <- TournDefense$'Pressures Succ'/TournDefense$'Pressures Press'
TournDefense$'Blocks Blocks/min' <- TournDefense$'Blocks Blocks'/TournDefense$'90s'
TournDefense$'Int/min' <- TournDefense$'Int'/TournDefense$'90s'
TournDefense$'Clr/min' <- TournDefense$'Clr'/TournDefense$'90s'
TournDefense$'Err/min' <- TournDefense$'Err'/TournDefense$'90s'
```

fix infinity & -infinity %
```{r}
LeagueShoot$'Gper90'[which(LeagueShoot$'Gper90' == Inf | LeagueShoot$'Gper90' == -Inf)] <- 0
TournShoot$'Gper90'[which(TournShoot$'Gper90' == Inf | TournShoot$'Gper90' == -Inf)] <- 0
LeagueShoot$'Performance%'[which(LeagueShoot$'Performance%' == Inf | LeagueShoot$'Performance%' == -Inf)] <- 0
TournShoot$'Performance%'[which(TournShoot$'Performance%' == Inf | TournShoot$'Performance%' == -Inf)] <- 0

LeagueDefense$'TklW%'[which(LeagueDefense$'TklW%' == Inf | LeagueDefense$'TklW%' == -Inf)] <- 0
TournDefense$'TklW%'[which(TournDefense$'TklW%' == Inf | TournDefense$'TklW%' == -Inf)] <- 0
LeagueDefense$'Tackles Tkl/min'[which(LeagueDefense$'Tackles Tkl/min' == Inf | LeagueDefense$'Tackles Tkl/min' == -Inf)] <- 0
TournDefense$'Tackles Tkl/min'[which(TournDefense$'Tackles Tkl/min' == Inf | TournDefense$'Tackles Tkl/min' == -Inf)] <- 0
LeagueDefense$'Vs Dribbles Past/min'[which(LeagueDefense$'Vs Dribbles Past/min' == Inf | LeagueDefense$'Vs Dribbles Past/min' == -Inf)] <- 0
TournDefense$'Vs Dribbles Past/min'[which(TournDefense$'Vs Dribbles Past/min' == Inf | TournDefense$'Vs Dribbles Past/min' == -Inf)] <- 0
LeagueDefense$'Pressures Succ %'[which(LeagueDefense$'Pressures Succ %' == Inf | LeagueDefense$'Pressures Succ %' == -Inf)] <- 0
TournDefense$'Pressures Succ %'[which(TournDefense$'Pressures Succ %' == Inf | TournDefense$'Pressures Succ %' == -Inf)] <- 0
LeagueDefense$'Blocks Blocks/min'[which(LeagueDefense$'Blocks Blocks/min' == Inf | LeagueDefense$'Blocks Blocks/min' == -Inf)] <- 0
TournDefense$'Blocks Blocks/min'[which(TournDefense$'Blocks Blocks/min' == Inf | TournDefense$'Blocks Blocks/min' == -Inf)] <- 0
LeagueDefense$'Int/min'[which(LeagueDefense$'Int/min' == Inf | LeagueDefense$'Int/min' == -Inf)] <- 0
TournDefense$'Int/min'[which(TournDefense$'Int/min' == Inf | TournDefense$'Int/min' == -Inf)] <- 0
LeagueDefense$'Clr/min'[which(LeagueDefense$'Clr/min' == Inf | LeagueDefense$'Clr/min' == -Inf)] <- 0
TournDefense$'Clr/min'[which(TournDefense$'Clr/min' == Inf | TournDefense$'Clr/min' == -Inf)] <- 0
LeagueDefense$'Err/min'[which(LeagueDefense$'Err/min' == Inf | LeagueDefense$'Err/min' == -Inf)] <- 0
TournDefense$'Err/min'[which(TournDefense$'Err/min' == Inf | TournDefense$'Err/min' == -Inf)] <- 0
```


#### Defender analysis ####

Selecting variants:

attributes: Pos, 90s, Tackles Tkl, Tackles TklW, Vs Dribbles Past, Pressures Succ, Pressures %, Blocks Blocks, Int, Clr, Err, TklW%

```{r}
defense.attributes <- c("Player", "90s", "Tackles Tkl/min", "TklW%", "Vs Dribbles Past/min", "Pressures Succ %", "Pressures %", "Blocks Blocks/min", "Int/min", "Clr/min", "Err/min")
LeagueDefense.data <- LeagueDefense[which(LeagueDefense$Pos != "GK"),defense.attributes]
TournDefense.data <- TournDefense[which(TournDefense$Pos != "GK"),c(defense.attributes, "Year")]
```

Fill in missing data

```{r}
for(attribute in defense.attributes){
  if(sum(is.na(LeagueDefense.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(LeagueDefense.data[,attribute])), "missing values in League Defense.", attribute))
    # fill the missing values with 0
    LeagueDefense.data[is.na(LeagueDefense.data[, attribute]), attribute] <- 0
  }
}
for(attribute in defense.attributes){
  if(sum(is.na(TournDefense.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(TournDefense.data[,attribute])), "missing values in Tournament Defense.", attribute))
    # fill the missing values with 0
    TournDefense.data[is.na(TournDefense.data[, attribute]), attribute] <- 0
  }
}
```


Group rows by player

```{r}
# Use the average value for each attribute
LeagueDefense.groupByPlayer <- LeagueDefense.data %>%
                                 group_by(Player) %>%
                                 summarise_all(mean)

#TournDefense.groupByPlayer <- TournDefense.data %>%
#                                 group_by(Player) %>%
#                                 summarise_all(mean)
```

#### Passing analysis ####

Selecting variants

attributes: 90s, xA, A-xA, KP, Total Cmp%
```{r}
pass.attributes <- c("Player","90s","xA","A-xA","KP","Total Cmp%")
LeaguePass.data <- LeaguePass[which(LeaguePass$Pos != "GK"), pass.attributes]
TournPass.data <- TournPass[which(TournPass$Pos != "GK"), c(pass.attributes, "Year")]
```

Fill in missing data

```{r}
# find out the columns with missing values
for(attribute in pass.attributes){
  if(sum(is.na(LeaguePass.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(LeaguePass.data[,attribute])), "missing values in League Passing.", attribute))
    # fill the missing values with 0
    LeaguePass.data[is.na(LeaguePass.data[, attribute]), attribute] <- 0
  }
}
for(attribute in pass.attributes){
  if(sum(is.na(TournPass.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(TournPass.data[,attribute])), "missing values in Tournament Passing.", attribute))
    # fill the missing values with 0
    TournPass.data[is.na(TournPass.data[, attribute]), attribute] <- 0
  }
}
```

Group rows by player

```{r}
# Use the average value for each attribute
library(dplyr)
LeaguePass.groupByPlayer <- LeaguePass.data %>%
                             group_by(Player) %>%
                               summarise_all(mean)

#TournPass.groupByPlayer <- TournPass.data %>%
#                                 group_by(Player) %>%
#                                 summarise_all(mean)
```

#### Shooting analysis ####

Selecting variants

attributes: Gper90, Standard SoT%, Standard G/Sh, Standard SoT/90, Performance%
```{r}
shoot.attributes <- c("Player", "90s", "Gper90", "Standard SoT%", "Standard G/Sh", "Standard SoT/90", "Performance%")
LeagueShoot.data <- LeagueShoot[which(LeagueShoot$Pos != "GK"), shoot.attributes]
TournShoot.data <- TournShoot[which(TournShoot$Pos != "GK"), c(shoot.attributes, "Year", "nation rank")]
#TournShoot.data <- TournShoot[which(TournShoot$Pos != "GK" & !is.na(TournShoot$`nation rank`)), c(shoot.attributes, "nation rank")]
```

```{r}
# find out the columns with missing values
for(attribute in shoot.attributes){
  if(sum(is.na(LeagueShoot.data[,attribute]))) {
    print(paste(sum(is.na(LeagueShoot.data[,attribute])), "missing values in League Forward.", attribute))
    LeagueShoot.data[is.na(LeagueShoot.data[, attribute]), attribute] <- 0
  }
}
for(attribute in shoot.attributes){
  if(sum(is.na(TournShoot.data[,attribute]))) {
    print(paste(sum(is.na(TournShoot.data[,attribute])), "missing values in Tournament Forward.", attribute))
    TournShoot.data[is.na(TournShoot.data[, attribute]), attribute] <- 0
  }
}
```


Group rows by player

```{r}
# Use the average value for each attribute
LeagueShoot.groupByPlayer <- LeagueShoot.data %>%
                                 group_by(Player) %>%
                                 summarise_all(mean)

#TournShoot.groupByPlayer <- TournShoot.data %>%
#                                 group_by(Player) %>%
#                                 summarise_all(mean)
```

Note: All players in tournament does not occur more than twice

```{r eval = F}
for(player in unique(TournDefense.data$Player)) {
  s <- length(which(TournDefense.data$Player == player))
  if(s > 2) print(s)
}
for(player in unique(TournPass.data$Player)) {
  s <- length(which(TournPass.data$Player == player))
  if(s > 2) print(s)
}
for(player in unique(TournShoot.data$Player)) {
  s <- length(which(TournShoot.data$Player == player))
  if(s > 2) print(s)
}
```

#### join tables
```{r}
League.tables <- list(LeagueShoot.groupByPlayer, LeaguePass.groupByPlayer, LeagueDefense.groupByPlayer)
Tourn.tables <- list(TournShoot.data, TournPass.data, TournDefense.data)

League.JoinedTable <- inner_join(inner_join(League.tables[[1]], League.tables[[2]], by="Player"), League.tables[[3]], by="Player")
Tourn.JoinedTable <- inner_join(inner_join(Tourn.tables[[1]], Tourn.tables[[2]], by=c("Player", "Year")), Tourn.tables[[3]], by=c("Player", "Year"))

# find the average playing time
League.JoinedTable$"90s" <- (League.JoinedTable$"90s" + League.JoinedTable$"90s.x" + League.JoinedTable$"90s.y")/3
Tourn.JoinedTable$"90s" <- (Tourn.JoinedTable$"90s" + Tourn.JoinedTable$"90s.x" + Tourn.JoinedTable$"90s.y")/3
League.JoinedTable <- League.JoinedTable[ , !(names(League.JoinedTable) %in% c("90s.x", "90s.y"))]
Tourn.JoinedTable <- Tourn.JoinedTable[ , !(names(Tourn.JoinedTable) %in% c("90s.x", "90s.y"))]
Tourn.JoinedTable.ranked <- Tourn.JoinedTable[which(!is.na(Tourn.JoinedTable$`nation rank`)),]
nrow(League.JoinedTable)
nrow(Tourn.JoinedTable)
nrow(Tourn.JoinedTable.ranked)
```


#### Write data to excel

```{r}
library("xlsx")
#write.xlsx(League.JoinedTable.Tailored, file="players_joined_subset.xlsx", sheetName = "League data", append = FALSE)
#write.xlsx(Tourn.JoinedTable.Tailored, file="players_joined_subset.xlsx", sheetName = "Tournament data", append = TRUE)
```

#### Write data to .rda

```{r}
save(
  LeagueShoot.data,
  LeagueShoot.groupByPlayer,
  LeagueDefense.data,
  LeagueDefense.groupByPlayer,
  LeaguePass.data,
  LeaguePass.groupByPlayer,
  
  TournShoot.data,
  TournDefense.data,
  TournPass.data,

  League.JoinedTable,
  Tourn.JoinedTable, 
  Tourn.JoinedTable.ranked,

  file = "players_joined_subset_not_grouped.rda")
```

```{r}
for(name in unique(LeagueShoot$Player)){
  if(length(unique(LeagueShoot[which(LeagueShoot$Player == name),]$Pos)) != 1) {
    print(name)
    print(unique(LeagueShoot[which(LeagueShoot$Player == name),]$Pos))
  }
}
```
