---
title: "Principle Component Analysis"
author: "Shuhao Liu"
date: "23/03/2022"
output: html_document
---
---
title: "knn all attributes"
output: html_document
---

```{r library}
library(readxl)
library(dplyr)
```

Load data
```{r}
file_path <- "/Users/liushuhao/Desktop/SOA Case Competitoin/data/2022-student-research-case-study-player-data_modified.xlsx"

#read in tables
LeagueShoot <- read_excel(file_path, sheet = "League Shooting", skip = 11,col_names =T, na = "N/A")
LeaguePass <- read_excel(file_path, sheet = "League Passing", skip = 11,col_names =T)
LeagueDefense <- read_excel(file_path, sheet = "League Defense", skip = 11,col_names =T, na = "N/A")
TournShoot <- read_excel(file_path, sheet = "Tournament Shooting", skip = 11,col_names =T, na = "N/A")
TournPass <- read_excel(file_path, sheet = "Tournament Passing", skip = 11,col_names =T)
TournDefense <- read_excel(file_path, sheet = "Tournament Defense", skip = 11,col_names =T, na = "N/A")
tables <- list("LeagueShoot" = LeagueShoot, "LeaguePass" = LeaguePass, "LeagueDefense" = LeagueDefense, "TournShoot" = TournShoot, "TournPass" = TournPass, "TournDefense" = TournDefense)
```


calculate %s
```{r eval=FALSE}
LeagueShoot$'Gls' <- LeagueShoot$'Gls'/LeagueShoot$'90s'
LeagueShoot$'Performance PKatt' <- LeagueShoot$'Performance PKatt'/LeagueShoot$'Performance PK'
TournShoot$'Gls' <- TournShoot$Gls/TournShoot$'90s'
TournShoot$'Performance PKatt' <- TournShoot$'Performance PKatt'/TournShoot$'Performance PK'

LeagueDefense$'Tackles TklW' <- LeagueDefense$'Tackles TklW'/LeagueDefense$'Tackles Tkl'
LeagueDefense$'Tackles Tkl' <- LeagueDefense$'Tackles Tkl'/LeagueDefense$'90s'
LeagueDefense$'Vs Dribbles Past' <- LeagueDefense$'Vs Dribbles Past'/LeagueDefense$'90s'
LeagueDefense$'Pressures Succ' <- LeagueDefense$'Pressures Succ'/LeagueDefense$'Pressures Press'
LeagueDefense$'Blocks Blocks' <- LeagueDefense$'Blocks Blocks'/LeagueDefense$'90s'
LeagueDefense$'Int' <- LeagueDefense$'Int'/LeagueDefense$'90s'
LeagueDefense$'Clr' <- LeagueDefense$'Clr'/LeagueDefense$'90s'
LeagueDefense$'Err' <- LeagueDefense$'Err'/LeagueDefense$'90s'
TournDefense$'Tackles TklW' <- TournDefense$'Tackles TklW'/TournDefense$'Tackles Tkl'
TournDefense$'Tackles Tkl' <- TournDefense$'Tackles Tkl'/TournDefense$'90s'
TournDefense$'Vs Dribbles Past' <- TournDefense$'Vs Dribbles Past'/TournDefense$'90s'
TournDefense$'Pressures Succ' <- TournDefense$'Pressures Succ'/TournDefense$'Pressures Press'
TournDefense$'Blocks Blocks' <- TournDefense$'Blocks Blocks'/TournDefense$'90s'
TournDefense$'Int' <- TournDefense$'Int'/TournDefense$'90s'
TournDefense$'Clr' <- TournDefense$'Clr'/TournDefense$'90s'
TournDefense$'Err' <- TournDefense$'Err'/TournDefense$'90s'
```

fix infinity & -infinity %
```{r eval=FALSE}
LeagueShoot$'Gls'[which(LeagueShoot$'Gls' == Inf | LeagueShoot$'Gls' == -Inf)] <- 0
TournShoot$'Gls'[which(TournShoot$'Gls' == Inf | TournShoot$'Gls' == -Inf)] <- 0
LeagueShoot$'Performance PKatt'[which(LeagueShoot$'Performance PKatt' == Inf | LeagueShoot$'Performance PKatt' == -Inf)] <- 0
TournShoot$'Performance PKatt'[which(TournShoot$'Performance PKatt' == Inf | TournShoot$'Performance PKatt' == -Inf)] <- 0

LeagueDefense$'Tackles TklW'[which(LeagueDefense$'Tackles TklW' == Inf | LeagueDefense$'Tackles TklW' == -Inf)] <- 0
TournDefense$'Tackles TklW'[which(TournDefense$'Tackles TklW' == Inf | TournDefense$'Tackles TklW' == -Inf)] <- 0
LeagueDefense$'Tackles Tkl'[which(LeagueDefense$'Tackles Tkl' == Inf | LeagueDefense$'Tackles Tkl' == -Inf)] <- 0
TournDefense$'Tackles Tkl'[which(TournDefense$'Tackles Tkl' == Inf | TournDefense$'Tackles Tkl' == -Inf)] <- 0
LeagueDefense$'Vs Dribbles Past'[which(LeagueDefense$'Vs Dribbles Past' == Inf | LeagueDefense$'Vs Dribbles Past' == -Inf)] <- 0
TournDefense$'Vs Dribbles Past'[which(TournDefense$'Vs Dribbles Past' == Inf | TournDefense$'Vs Dribbles Past' == -Inf)] <- 0
LeagueDefense$'Pressures Succ'[which(LeagueDefense$'Pressures Succ' == Inf | LeagueDefense$'Pressures Succ' == -Inf)] <- 0
TournDefense$'Pressures Succ'[which(TournDefense$'Pressures Succ' == Inf | TournDefense$'Pressures Succ' == -Inf)] <- 0
LeagueDefense$'Blocks Blocks'[which(LeagueDefense$'Blocks Blocks' == Inf | LeagueDefense$'Blocks Blocks' == -Inf)] <- 0
TournDefense$'Blocks Blocks'[which(TournDefense$'Blocks Blocks' == Inf | TournDefense$'Blocks Blocks' == -Inf)] <- 0
LeagueDefense$'Int'[which(LeagueDefense$'Int' == Inf | LeagueDefense$'Int' == -Inf)] <- 0
TournDefense$'Int'[which(TournDefense$'Int' == Inf | TournDefense$'Int' == -Inf)] <- 0
LeagueDefense$'Clr'[which(LeagueDefense$'Clr' == Inf | LeagueDefense$'Clr' == -Inf)] <- 0
TournDefense$'Clr'[which(TournDefense$'Clr' == Inf | TournDefense$'Clr' == -Inf)] <- 0
LeagueDefense$'Err'[which(LeagueDefense$'Err' == Inf | LeagueDefense$'Err' == -Inf)] <- 0
TournDefense$'Err'[which(TournDefense$'Err' == Inf | TournDefense$'Err' == -Inf)] <- 0
```

#### Defender analysis ####

```{r}
LDefense.NAattributes <- c(2, 3, 4, 6, 31, 32)
TDefense.NAattributes <- c(2, 3, 5, 30, 31)
LeagueDefense.data <- LeagueDefense[which(LeagueDefense$Pos != "GK"), -LDefense.NAattributes]
TournDefense.data <- TournDefense[which(TournDefense$Pos != "GK"), -TDefense.NAattributes]
```

Fill in missing data

```{r}
for(attribute in 1:ncol(LeagueDefense.data)){
  if(sum(is.na(LeagueDefense.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(LeagueDefense.data[,attribute])), "missing values in League Defense.", attribute))
    # fill the missing values with 0
    LeagueDefense.data[is.na(LeagueDefense.data[, attribute]), attribute] <- 0
  }
}
for(attribute in 1:ncol(TournDefense.data)){
  if(sum(is.na(TournDefense.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(TournDefense.data[,attribute])), "missing values in Tournament Defense.", attribute))
    # fill the missing values with 0
    TournDefense.data[is.na(TournDefense.data[, attribute]), attribute] <- 0
  }
}
```


Group rows by player

```{r}
# Use the average value for each attribute
LeagueDefense.groupByPlayer <- LeagueDefense.data %>%
                                 group_by(Player) %>%
                                 summarise_all(mean)

TournDefense.groupByPlayer <- TournDefense.data %>%
                                 group_by(Player) %>%
                                 summarise_all(mean)
```

#### Passing analysis ####

Selecting variants

attributes: 90s, xA, A-xA, KP, Total Cmp%
```{r}
LPass.NAattributes <- c(2, 3, 4, 6, 30, 31)
TPass.NAattributes <- c(2, 3, 5, 29, 30)
LeaguePass.data <- LeaguePass[which(LeaguePass$Pos != "GK"), -LPass.NAattributes]
TournPass.data <- TournPass[which(TournPass$Pos != "GK"), -TPass.NAattributes]
```

Fill in missing data

```{r}
# find out the columns with missing values
for(attribute in 1:ncol(LeaguePass.data)){
  if(sum(is.na(LeaguePass.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(LeaguePass.data[,attribute])), "missing values in League Passing.", attribute))
    # fill the missing values with 0
    LeaguePass.data[is.na(LeaguePass.data[, attribute]), attribute] <- 0
  }
}
for(attribute in 1:ncol(TournPass.data)){
  if(sum(is.na(TournPass.data[,attribute]))) {
    # find out the columns with missing values
    print(paste(sum(is.na(TournPass.data[,attribute])), "missing values in Tournament Passing.", attribute))
    # fill the missing values with 0
    TournPass.data[is.na(TournPass.data[, attribute]), attribute] <- 0
  }
}
```

Group rows by player

```{r}
# Use the average value for each attribute
library(dplyr)
LeaguePass.groupByPlayer <- LeaguePass.data %>%
                              group_by(Player) %>%
                              summarise_all(mean)

TournPass.groupByPlayer <- TournPass.data %>%
                             group_by(Player) %>%
                             summarise_all(mean)
```


#### Shooting analysis ####

Selecting variants

```{r}
LShoot.NAattributes <- c(2, 3, 4, 6, 17, 20, 21, 22, 23, 24, 25, 26)
TShoot.NAattributes <- c(2, 3, 5, 18, 19)
LeagueShoot.data <- LeagueShoot[which(LeagueShoot$Pos != "GK"), -LShoot.NAattributes]
TournShoot.data <- TournShoot[which(TournShoot$Pos != "GK"), -TShoot.NAattributes]
```

```{r}
# find out the columns with missing values
for(attribute in 1:ncol(LeagueShoot.data)){
  if(sum(is.na(LeagueShoot.data[,attribute]))) {
    print(paste(sum(is.na(LeagueShoot.data[,attribute])), "missing values in League Forward.", attribute))
    LeagueShoot.data[is.na(LeagueShoot.data[, attribute]), attribute] <- 0
  }
}
for(attribute in 1:ncol(TournShoot.data)){
  if(sum(is.na(TournShoot.data[,attribute]))) {
    print(paste(sum(is.na(TournShoot.data[,attribute])), "missing values in Tournament Forward.", attribute))
    TournShoot.data[is.na(TournShoot.data[, attribute]), attribute] <- 0
  }
}
```


Group rows by player

```{r}
# Use the average value for each attribute
LeagueShoot.groupByPlayer <- LeagueShoot.data %>%
                                 group_by(Player) %>%
                                 summarise_all(mean)

TournShoot.groupByPlayer <- TournShoot.data %>%
                                 group_by(Player) %>%
                                 summarise_all(mean)
```

Note: All players in tournament does not occur more than twice

```{r eval = F}
for(player in unique(TournDefense.data$Player)) {
  s <- length(which(TournDefense.data$Player == player))
  if(s > 2) print(s)
}
for(player in unique(TournPass.data$Player)) {
  s <- length(which(TournPass.data$Player == player))
  if(s > 2) print(s)
}
for(player in unique(TournShoot.data$Player)) {
  s <- length(which(TournShoot.data$Player == player))
  if(s > 2) print(s)
}
```

#### join tables
```{r}
tables.gbp <- list("ls.gbp" = LeagueShoot.groupByPlayer, "ts.gbp" = TournShoot.groupByPlayer, "ld,gbp" = LeagueDefense.groupByPlayer, "td.gbp" = TournDefense.groupByPlayer, "lp.gbp" = LeaguePass.groupByPlayer, "tp.gbp" = TournPass.groupByPlayer)
League.tables <- list(LeagueShoot.groupByPlayer, LeaguePass.groupByPlayer, LeagueDefense.groupByPlayer)
Tourn.tables <- list(TournShoot.groupByPlayer, TournPass.groupByPlayer, TournDefense.groupByPlayer)

League.JoinedTable <- inner_join(inner_join(League.tables[[1]], League.tables[[2]], by="Player"), League.tables[[3]], by="Player")
Tourn.JoinedTable <- inner_join(inner_join(Tourn.tables[[1]], Tourn.tables[[2]], by="Player"), Tourn.tables[[3]], by="Player")

# find the average playing time
League.JoinedTable$"90s" <- (League.JoinedTable$"90s" + League.JoinedTable$"90s.x" + League.JoinedTable$"90s.y")/3
Tourn.JoinedTable$"90s" <- (Tourn.JoinedTable$"90s" + Tourn.JoinedTable$"90s.x" + Tourn.JoinedTable$"90s.y")/3
League.JoinedTable <- League.JoinedTable[ , !(names(League.JoinedTable) %in% c("90s.x", "90s.y"))]
Tourn.JoinedTable <- Tourn.JoinedTable[ , !(names(Tourn.JoinedTable) %in% c("90s.x", "90s.y"))]
Tourn.JoinedTable.ranked <- Tourn.JoinedTable[which(!is.na(Tourn.JoinedTable$`nation rank`)),]
nrow(League.JoinedTable)
nrow(Tourn.JoinedTable)
```

Normalize Data

```{r}
#the normalization function is created
norm <-function(x) { 
  upper = quantile(x, .98)
  lower = quantile(x, .02)
  x_norm <- (x-lower)/(upper-lower)
  x_norm[which(x_norm > 1)] = 1
  x_norm[which(x_norm < 0)] = 0
  x_norm
}

playerIndex <- which(colnames(Tourn.JoinedTable.ranked) == "Player")
rankIndex <- which(colnames(Tourn.JoinedTable.ranked) == "nation rank")
Tourn_norm <- as.data.frame(lapply(Tourn.JoinedTable.ranked[,-c(playerIndex,rankIndex)], norm))

LeagePlayerIndex <- which(colnames(League.JoinedTable) == "Player")
League_norm <- as.data.frame(lapply(League.JoinedTable[,-LeagePlayerIndex], norm))
```

#### Principle Component Analysis
```{r fig.width = 10, fig.height= 4}
# decide how many components to use
eigenvalues <- eigen(cor(League_norm))$values
eigenvaluesProp <- cumsum(eigenvalues)/sum(eigenvalues)
par(mfrow = c(1,2), cex = 0.7)
plot(eigenvalues, main = "Eigenvalues of the Correlation Matrix", sub = "Figure 2.1 Eigenvalues of the Correlation Matrix")
plot(eigenvaluesProp, xlab = "number of components", ylab = "proportion", main = "Proportion of Variation Included", sub = "Figure 2.2 Variation explained corresponding to the number of components used in PCA")
abline(h = 0.9, lty = 2, col = "red")
abline(v = 21, lty = 21, col = "red")
points(21, eigenvaluesProp[21], pch = 19, col = "red", cex = 0.95)
text(26, 0.85, col = "red", "(21, 0.9)")
```
We'll choose 26 components to over 90% of the variation. 

```{r}
#check if the variance of different attributes are large
apply(Tourn_norm, 2, var)
plot(apply(Tourn_norm, 2, var))
```

```{r}
# apply Principle Component Analysis
compNum <- 22
pca = prcomp(Tourn_norm, scale = T)
Tourn_norm.pca <- as.matrix(Tourn_norm) %*% as.matrix(pca$rotation[,1:compNum])
```


```{r}
#run knn function
library(class)
k <- 18
error <- 9
A <- c()
B <- c()
C <- c()
for(i in 1:2000) {
TournRanks <- Tourn.JoinedTable.ranked$`nation rank`
pr <- knn.cv(Tourn_norm.pca,cl=TournRanks,k=k,use.all = TRUE)
accuracy9 <- sum(abs(as.numeric(pr) - as.numeric(TournRanks)) <= error) / length(pr)
accuracy0 <- sum(as.numeric(pr) == as.numeric(TournRanks)) / length(pr)
accuracy1 <- sum(abs(as.numeric(pr) - as.numeric(TournRanks)) <= 1) / length(pr)
A = c(A, accuracy9)
B <- c(B, accuracy0)
C <- c(C, accuracy1)
}
mean(A)
mean(B)
mean(C)
```


```{r fig.width=10, fig.height=4}
par(mfrow = c(1,2), cex = 0.8)
hist(A+0.001, breaks = 25, probability = T, xlab = "probability of ranked top 10", main = "Histogram of the Probability of Ranked Top 10", sub = "Figure 2.3 Histogram of the Probability of Reaching Top 10")
hist(C, breaks = 25, probability = T, xlab = "probability of ranked top 10", main = "Histogram of the Probability of Achieving Championship", sub = "Figure 2.4 Histogram of the Probability of Achieving Championship")
```
```{r}
#hist(B, breaks = 12, probability = T, xlab = "probability of ranked top 10", main = "Histogram of the Probability Reaching Top 10")
hist(C, breaks = 25, probability = T, xlab = "probability of ranked top 10", main = "Histogram of the Probability Reaching Top 1")
```

```{r}
mean(A) - sd(A)*1.96
mean(A) + sd(A)*1.96
```

```{r}
mean(C) - sd(C)*1.96
mean(C) + sd(C)*1.96
```

```{r}
League_norm.pca <- as.matrix(League_norm) %*% as.matrix(pca$rotation[,1:compNum])
Target_category <- as.data.frame(Tourn.JoinedTable.ranked)[,rankIndex]
predRank <- knn(Tourn_norm.pca,League_norm.pca,cl=Target_category,k=k)
#League.JoinedTable$"Pred Rank" <- predRank
table(predRank)
```
